---
title: "p8105_hw2_cz2657"
output: github_document
---

## Problem 1

Reading and cleaning Mr. Trash Wheel sheet

```{r, message=FALSE}
library(tidyverse)
library(readxl)
```


```{r, message = FALSE}
trash_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  drop_na("dumpster") %>% 
  mutate(sports_balls = round(as.numeric(sports_balls)))

rain2018_df = read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14", col_names = TRUE) %>%
  mutate(year = 2018) %>% 
  janitor::clean_names() %>% 
  drop_na(total)

rain2019_df = read_excel("data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", range = "A2:B14", col_names = TRUE) %>%
  mutate(year = 2019) %>% 
  janitor::clean_names() %>% 
  drop_na(total)

rain_df = full_join(rain2018_df, rain2019_df) %>% 
  mutate(month = month.name[month]) %>% 
  rename(total_precipitation_in = total)

```

The `trash_df` dataset has `r nrow(trash_df)` observations and includes the variables `weight_tons` and `volume_cubic_yards` which describe the amount of garbage consumed by Mr. Trash Wheel. Furthermore, it provides detail around the number of plastic bottles (`plastic_bottles`), cigarette butts (`cigarette_butts`), and grocery bags (`grocery_bags`) consumed by Mr. Trash Wheel. Excitingly, Mr. Trash Wheel powered up to `r max(pull(trash_df, homes_powered), na.rm=TRUE)` homes using the trash collected from a single dumpster. The median number of `sports_balls` in a dumpster in 2019 was `r median(pull(filter(trash_df, year == 2019),sports_balls))`. 

The `rain_df`dataset has `r nrow(rain_df)` observations and includes the variables `r colnames(rain_df)`. Total precipitation in 2018 was `r sum(pull(filter(rain_df, year == 2018), total_precipitation_in))` inches. 

## Problem 2 

```{r, message = FALSE}
pols_df = read_csv("data/pols-month.csv") %>% 
  janitor::clean_names() %>% 
  separate(mon, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(month = month.name[month], president = ifelse(prez_gop, "gop", "dem")) %>% 
  select (-prez_dem, -prez_gop, -day)

snp_df = read_csv("data/snp.csv") %>% 
  janitor::clean_names() %>% 
  mutate(date = as.Date(date,"%m/%d/%y ")) %>% 
  separate(date, into = c("year", "month", "day"), convert = TRUE) %>% 
  mutate(month = month.name[month]) %>% 
  select(year, month, snp_close_value = close, -day)

unemployment_df = read_csv("data/unemployment.csv") %>% 
  pivot_longer(Jan:Dec, names_to = "month") %>% 
  mutate(month = match(month, month.abb), month = month.name[month]) %>% 
  janitor::clean_names() %>% 
  rename(percent_unemployed = value)

fte_df = left_join(pols_df, snp_df, by = c("year", "month")) %>% 
  left_join(unemployment_df, by = c("year", "month"))

```

The `prez_df` dataset provides information on the number of Republican and Democratic government officials during each month beginning in 1947, including senators (`sen_gop` and `sen_dem`), representatives (`rep_gop` and `rep_dem`), and governors (`gov_gop` and `gov_dem`), as well as the party of the `president`. It has `r ncol(pols_df)`variables and `r nrow(pols_df)` observations. 

The `snp_df` dataset provides the closing value of the S&P stock index on a given date of each month in a given year. It has `r ncol(snp_df)` variables and `r nrow(snp_df)` observations. 

The `unemployment_df` dataset provides information on the percentage of unemployment in a given `month` of a given `year`. It has `r ncol(unemployment_df)` variables and `r nrow(unemployment_df)` observations.

Finally, the compiled dataset, `fte_df` combines this information by month and year to provide the unemployment percentage, S&P closing value, and political representation information on a monthly basis from ` r min(pull(fte_df, year))` to `r max(pull(fte_df, year))`, although some datapoints are missing especially in earlier years. The dataset has `r nrow(fte_df)` observations and `r ncol(fte_df)` variables, including  all `president`, `percent_unemployed`, and `snp_close_value`. 

## Problem 3

Loading and tidying the baby names data

```{r, message = FALSE}
baby_df = read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names() %>% 
  mutate(ethnicity = str_to_lower(ethnicity), gender = str_to_lower(gender), ethnicity = recode (ethnicity, "asian and paci" = "asian and pacific islander", "black non hisp" = "black non hispanic", "white non hisp" = "white non hispanic")) %>% 
  distinct(year_of_birth, gender, ethnicity, childs_first_name, count, rank, .keep_all = TRUE) 

```


```{r}

olivia_df = baby_df %>% 
  select (-count) %>% 
  pivot_wider(names_from = "year_of_birth", values_from = "rank") %>% 
  filter(childs_first_name == "Olivia" & gender == "female") %>% 
  select(-gender, - childs_first_name) %>% 
  knitr::kable(caption = "Popularity of the name Olivia as a female baby name over time") 

male_df = baby_df %>% 
  select (-count) %>% 
  filter(gender == "male", rank == 1) %>% 
  pivot_wider(names_from = "year_of_birth", values_from = "childs_first_name") %>% 
  select(-gender, -rank) %>% 
  knitr::kable(caption = "Most popular name among male children over time") 

names_plot = baby_df %>% 
  filter(gender == "male", ethnicity == "white non hispanic", year_of_birth == 2016) %>% 
  ggplot(aes(x = rank, y = count)) +
  geom_point(color = "blue", alpha = .3) +
  labs(title = "Count of Children with a given name by Rank")

names_plot
```


